---
title: "Problem Set 1"
author: "Joe Ornstein"
date: "August 16, 2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
library(tidyverse)
```

## Introducing RMarkdown

This is an RMarkdown document, which blends text and code into a single document. It's great for **reproducible research**, because it provides a complete step-by-step record of everything you did to generate your results!

For example, the following code chunk will produce a chart of socioeconomic indicators from 1888 in forty-seven Swiss provinces.

```{r swiss-plot, echo=FALSE}
# this creates a ggplot() object, using the swiss dataset that comes pre-packaged with R
ggplot(data = swiss) +
  # add points, where x-axis is mapped to Agriculture and y-axis is mapped to Education 
  geom_point(mapping = aes(x = Agriculture, 
                           y = Education),
             alpha = 0.7) +
  # black and white theme
  theme_bw() +
  # axis labels
  labs(x = 'Percent of Males In Agricultural Occupations',
       y = 'Percent Education Beyond Primary School')
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing the R code that generated the plot.

## References
