---
title: "Perceptions of Corruption by Income Level"
author: "Ansley Hatcher"
date: "9/20/2021"
output: pdf_document
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

Question Q112 on the World Values Survey Wave 7 asks respondents to record "[their] views on corruption - when people pay a bribe, give a gift, or do a favor to other people in order to get the things they need done or the services they need" @haerpfer2020. Respondents placed their views on corruption in their country on a 10-point scale, "where '1' means 'there is no corruption in this country' and '10' means 'there is abundant corruption in this country.'" The survey also has respondents record what income group their household is in by specifying the number of all wages, salaries, pensions, and other income the household collects. The surveyors organized income level responses into three categories: low, middle, and high. I compared respondents' answers to these two questions and compared trends in responses between countries.

## Figure 1

```{r pressure7, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(here)
library(forcats)
load(here('data/raw/WVS_Cross-National_Wave_7_R_v2_0.rdata'))

# country: C_COW_ALPHA
# perceptions of corruption in the country: Q112 
  # Now I'd like you to tell me your views on corruption – when people pay a bribe, give a
  # gift or do a favor to other people in order to get the things they need done or the services
  # they need. How would you place your views on corruption in your country on a 10-point
  # scale where “1” means “there is no corruption in this country” and “10” means “there
  # is abundant corruption in this country”. If your views are somewhat mixed, choose the
  # appropriate number in between.
# Income level (recoded): Q288R
  # Income level recoded: Low - steps 1-3, Middle - steps 4-7, High - steps 8-10
  # 1 = Low, 2 = Medium, 3 = High

x <- `WVS_Cross-National_Wave_7_v2_0` %>%
  select(country = C_COW_ALPHA,
         corruption_country = Q112,
         income_level = Q288R) %>%
  group_by(income_level, country) %>%
  summarize(mean_corruption_country = mean(corruption_country),
            count = n()) %>%
  filter(country %in% c('USA', 'CHN', 'JPN', 'ROK', 'BRA', 'NEW'),
        income_level %in% c(1, 2, 3)) %>%
  mutate(country = recode(country, "BRA" = "Brazil", "CHN" = "China", "JPN" = "Japan", "NEW" = "New Zealand", "ROK" = "South Korea", "USA" = "United States")) %>%
  ggplot(mapping = aes(x=income_level, y = mean_corruption_country,
                       group = country)) +
  geom_line() +
  theme_bw() +
  facet_wrap(~country) +
  labs(x = 'Income Level (Low to High)', y = 'Mean Perception of Corruption in the Country', caption = 'Data source: World Values Survey Wave 7', title = 'Mean Perception of Corruption by Income Level') +
  scale_y_continuous(limits = c(0, 10))
x
```

There is a clear trend in the United States where as income decreases, perception of corruption in the country increases. There is a similar trend in Figure 1 for Brazil and New Zealand. In Brazil, this trend is much less significant as perceptions of corruption are high overall. Next, I broke down perceptions of corruption further by examining perceptions of state authorities and business executives.

## Figure 2

```{r pressure9, echo=FALSE, message=FALSE, warning=FALSE}
# Involved in corruption: State authorities: Q113
  # Among the following groups of people, how many do you believe are involved in
  # corruption? Tell me for each group if you believe it is none of them, few of them, most
  # of them or all of them? State authorities
# Income level (recoded): Q288R
  # Income level recoded: Low - steps 1-3, Middle - steps 4-7, High - steps 8-10
  # 1 = Low, 2 = Medium, 3 = High
# country: C_COW_ALPHA

p <- `WVS_Cross-National_Wave_7_v2_0` %>%
  select(country = C_COW_ALPHA,
         corruption_stateauthorities = Q113,
         income_level = Q288R) %>%
  group_by(income_level, country) %>%
  summarize(mean_corruption_stateauthorities = mean(corruption_stateauthorities),
            count = n()) %>%
  filter(country %in% c('USA', 'CHN', 'JPN', 'ROK', 'BRA', 'NEW'),
         income_level %in% c(1, 2, 3)) %>%
  mutate(country = recode(country, "BRA" = "Brazil", "CHN" = "China", "JPN" = "Japan", "NEW" = "New Zealand", "ROK" = "South Korea", "USA" = "United States")) %>%
  ggplot(mapping = aes(x = income_level, y = mean_corruption_stateauthorities,
                       group = country)) +
  geom_line() +
  theme_bw() +
  facet_wrap(~country) +
  labs(x = 'Income Level (Low to High)', y = 'Mean Perception of Corruption of State Authorities', caption = 'Data source: World Values Survey Wave 7', title = 'Mean Perception of Corruption of State Authorities by Income Level') +
   scale_y_continuous(limits = c(0, 4))
p
```

The survey breaks down perceptions of corruption by asking respondents "Among the following groups of people, how many do you believe are involved in corruption?" @haerpfer2020 The respondents reported their feelings on a 1-4 scale, where 1 is "None of them," 2 is "Few of them," 3 is "Most of them," and 4 is "All of them."

The trend in the United States in Figure 2 is similar to Figure 1, where perceptions of corruption are highest among those of lower income. Interestingly, the trend flipped in New Zealand, where those of higher income perceive more corruption among state officials than those of lower income.

## Figure 3

```{r pressure13, echo=FALSE, message=FALSE, warning=FALSE}
# Involved in corruption: Business executives: Q114
 #Among the following groups of people, how many do you believe are involved in
  # corruption? Tell me for each group if you believe it is none of them, few of them, most
  # of them or all of them? Business executives
# country: C_COW_ALPHA
# Income level (recoded): Q288R
  # Income level recoded: Low - steps 1-3, Middle - steps 4-7, High - steps 8-10
  # 1 = Low, 2 = Medium, 3 = High
p <- `WVS_Cross-National_Wave_7_v2_0` %>%
  select(country = C_COW_ALPHA,
         corruption_businessexecutives = Q114,
         income_level = Q288R) %>%
  group_by(income_level, country) %>%
  summarize(mean_corruption_businessexecutives = mean(corruption_businessexecutives),
            count = n()) %>%
  filter(country %in% c('USA', 'CHN', 'JPN', 'ROK', 'BRA', 'NEW'),
         income_level %in% c(1, 2, 3)) %>%
  mutate(country = recode(country, "BRA" = "Brazil", "CHN" = "China", "JPN" = "Japan", "NEW" = "New Zealand", "ROK" = "South Korea", "USA" = "United States")) %>%
  ggplot(mapping = aes(x = income_level, y = mean_corruption_businessexecutives,
                       group = country)) +
  geom_line() +
  theme_bw() +
  facet_wrap(~country) +
  labs(x = 'Income Level (Low to High)', y = 'Mean Perception of Corruption of Business Executives', caption = 'Data source: World Values Survey Wave 7', title = 'Mean Perception of Corruption of Business Executives by Income Level') +
   scale_y_continuous(limits = c(0, 4))
p
```

One last question from the survey I chose to examine has respondents record their perceived level of corruption of business executives, using the same 1-4 scale. Not surprisingly, the trends stayed mostly the same. In South Korea, there is a sharp decline in perceived corruption by those of a higher income, probably because they are the ones who would be defined as business executives.

# Results

For several countries, there is a clear trend of higher perceived corruption among those of lower income. This is probably due to the fact that most of those with lower income are more likely to be affected by corruption. Additionally, those with higher income are more likely to commit acts of corruption and are not going to reflect that on a survey. Previous research has found similar results, that "those most harmed by corruption - the socioeconomically disadvantaged - should perceive corruption to be more frequent" @maeda2015.

# References
