---
title: "Compulsory Voting in Karnataka, India"
author: "Reshi Rajan"
date: "21 September 2021"
output:
  pdf_document: default
  word_document: default
  html_document:
    df_print: paged
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Compulsory Voting in the 2015 Karnataka Gram Panchayat Election

In 2015, the south-west Indian state of Karnataka implemented compulsory voting in their local gram, taluk, and zilla panchayat elections ([@aji2015]). In addition, in 2009, the state of Gujarat in north-west India attempted to implement compulsory voting, though the Gujarat High Court ruled this to be illegal and compulsory voting did not occur in Gujarat ([@pti]). The use of compulsory voting in local body elections has been advocated by Indian politicians who believe turnout is significantly lower in these elections compared to legislative assembly elections and Indian general elections, and more voters should be informed of their voting rights ([@khan2015]).

Based on the work of Singh 2010, compulsory voting even without penalties should contribute to an increase in voter turnout. Thus, it would be reasonable to assume that the 2015 Karnataka gram panchayat elections would see voter turnout increase. Some studies that use aggregate data suggest the force of abstention penalties effect turnout, but do not include other factors linked to turnout, such as age, education, or income, which when disaggregated show that compulsory voting without penalties can boost turnout ([@singh2011]). Furthermore, the work of Green and Shachar 2000, suggests that voting patterns in one election influence the voting patterns of the next, thus despite the removal of compulsory voting in the 2020, it should be expected that voter turnout will still be higher relative to the pre-2015 period.

Thus, we postulate:

(H1): Compulsory voting increased turnout in Karnataka during the 2015 gram panchayat election.

In order to determine whether voter turnout increased meaningfully in Karnataka, a difference-in-difference analysis was conducted. To compare results with Karnataka, the states of Kerala, Rajasthan, Tamil Nadu, and West Bengal were chosen as control states. All states had a similar organizational structure of panchayats when compared to each other and Karnataka.

In addition, states were chosen based on population size, political trends, geographic location, human development levels and synchronized election cycles. Rajasthan and Kerala were both chosen as their panchayat elections are conducted in the same years as Karnataka, allowing for a better understanding of any national trends within India. In addition, Rajasthan was chosen because of its similar political trends, i.e., the two major parties in both Rajasthan and Karnataka are the BJP and INC. Kerala and Tamil Nadu were both chosen as they are geographically close to Karnataka, being its southern and south-eastern neighbor, respectively. In addition, Kerala, Tamil Nadu, and Karnataka are all Dravidian-speaking areas with significant cultural similarities and levels of human development. West Bengal and Rajasthan were chosen for their similar population size to Karnataka.

## Results

When plotting the data from all available time periods in the state of Karnataka and the control states of Rajasthan and Kerala. The following plots are generated.

```{r, echo=FALSE, warning=FALSE, message=FALSE}
library(tidyverse) 
load("D:/RStudio/POLS 7012/intro-political-methodology/data/CV Data.RData")

q <- `CV` %>%
  group_by(State) %>%
  filter(State %in% c('Kerala')) %>%
  ggplot(mapping = aes(x= `State`, y= `Turnout`, color=`Turnout`,
                       group = `State`, `District`)) +
  facet_wrap(~Year)+
  geom_point() +
  theme_minimal()+
  labs(x = 'Districts by State - Kerala', y = 'Turnout')
q + theme_minimal()
```

```{r, echo=FALSE}
x <- `CV` %>%
  group_by(State) %>%
  filter(State %in% c('Karnataka')) %>%
  ggplot(mapping = aes(x= `State`, y= `Turnout`, color=`Turnout`,
                       group = `State`, `District`)) +
  facet_wrap(~Year)+
  geom_point() +
  theme_minimal()+
  labs(x = 'Districts by State - Karnataka', y = 'Turnout')
x + theme_minimal()
```

```{r, echo=FALSE}
y <- `CV` %>%
  group_by(State) %>%
  filter(State %in% c('Rajasthan')) %>%
  ggplot(mapping = aes(x= `State`, y= `Turnout`, color=`Turnout`,
                       group = `State`,`District`)) +
  facet_wrap(~Year)+
  geom_point() +
  theme_minimal()+
  labs(x = 'Districts by State - Rajasthan', y = 'Turnout')
y + theme_minimal()

```

## Difference-in-Difference Analysis

The results of the difference-in-difference analysis are shown below. The tables show how data from Karnataka compared to the data of the control states of Kerala and Rajasthan. As seen in Figure 1, the treated group (i.e., Karnataka in 2015) had a significant increase in voter turnout.

```{r, echo=FALSE}

CV$Time = ifelse(CV$Year >= 2015, 1, 0)
CV$Treated = ifelse(CV$State == 'Karnataka', 1, 0)
CV$did = CV$Time * CV$Treated 
didreg = lm(CV$Turnout ~ Treated + Time + did, data = CV)
summary(didreg)

ggplot(CV,aes(x=Year,y=Turnout,colour=factor(did)))+geom_point()+
stat_smooth(method='lm',formula=y~scale(x,scale=F))+ggtitle('Figure 1. Regression Analysis of Difference-in-Difference Analysis')

```

We then ran the difference in difference analysis without running the interaction. Even without the interaction between the variables generated, the data shows that voter turnout still increased meaningfully. The results of this data is shown below,

```{r, echo=FALSE}

didreg1 = lm (CV$Turnout ~ Treated*Time, data = CV)
summary(didreg1)

```

## Analysis

The difference-in-difference analysis presents evidence showing that compulsory voting did cause voter turnout to increase in the 2015 Karnataka gram panchayat election. These results further contribute to the existing literature showing that compulsory voting laws, even without penalties for abstention, do increase voter turnout ([@singh2011]). These results have implications for the use of compulsory voting laws in the future. With India having generally high voter turnout for a voluntary voting system ([@banerjee2020]), the implementation of compulsory voting without penalties shows that voter turnout can still increase without the need to penalize voters for abstaining. This could provide an option for governments who wish to increase turnout without penalizing voters for abstention.

## Bibliography
