---
title: "Integral Calculus"
#author: "Joe Ornstein"
#date: "September 30, 2020"
output:
  xaringan::moon_reader:
    nature:
      highlightStyle: github
      countIncrementalSlides: false
# subtitle: 'POLS 7012: Introduction to Political Methodology'
#institute: University of Georgia
---

```{r Setup, include=FALSE}
library(tidyverse)
theme_set(theme_minimal(base_size = 16))
knitr::opts_chunk$set(comment=NA, fig.width=7, fig.height=5, 
                      fig.align = 'center', out.width = 600,
                      message=FALSE, warning=FALSE, echo=FALSE)
```

# The Problem

To compute a p-value, we need to know the area of the shaded part of the probability distribution function.

```{r, echo = FALSE, out.width='400px'}

# plot a gaussian and shade a part of it
mu <- 0
sigma <- 1

# http://www.sthda.com/english/wiki/ggplot2-area-plot-quick-start-guide-r-software-and-data-visualization
d <- tibble(x = seq(-4, 4, 0.01),
            y = dnorm(x, mean = mu, sd = sigma))

ggplot(data = d) +
  geom_line(mapping = aes(x=x,y=y)) +
  geom_area(mapping = aes(x=x, y = if_else(x > 2, y, 0)),
            fill = 'gray', alpha = 0.5) + 
  theme_void() +
  geom_vline(xintercept = mu, linetype = 'dashed')
```


Formally:

$$\int_{\hat{\tau}}^\infty \mathcal{N}\left(0, \frac{\sigma^2}{n}\right)$$

---

# The Area Under a Curve

Rectangles are easy. $A = bh$

```{r rectangular area}

data = tibble(x = -2:2,
              y = 2)

ggplot() +
  geom_rect(aes(xmin = -2, xmax = 2, ymin = 0, ymax = 2),
            alpha = 0.3, fill = 'blue') + 
  geom_line(data = data, aes(x=x,y=y), size = 1) +
  scale_x_continuous(limits = c(-4,4)) +
  scale_y_continuous(limits = c(-4,4)) +
  geom_hline(yintercept = 0) + geom_vline(xintercept = 0) + 
  xlab('x') + ylab('y') + 
  geom_text(aes(x=2.4,y=2, label='f(x)'), size = 5)


```

---

# The Area Under a Curve

Curves are harder. We didn't learn this in geometry!


```{r plot curve}

f <- function(x){
  x^3 - 2*x^2 + 2
}

# f <- function(x){
#   0.7*x^2
# }

x <- seq(-2,2,0.01)

data <- tibble(x=x,
               y=f(x))

ggplot() +
  geom_line(data = data, aes(x=x,y=y), size = 1) +
  scale_x_continuous(limits = c(-4,4)) +
  scale_y_continuous(limits = c(-4,4)) +
  geom_hline(yintercept = 0) + geom_vline(xintercept = 0) + 
  xlab('x') + ylab('y') + 
  geom_text(aes(x=2.4,y=2, label='f(x)'), size = 5)


```

---

# The Area Under a Curve

 - With derivatives, we approximated a hard problem (the slope of a curve) using an easy problem (the slope of a line) by zooming in close enough.

--

 - With integrals, we'll use a similar trick.

--

 - We approximate a hard problem (the area under a curve) using an easy problem (the area of a bunch of rectangles) by making the rectangles *really* thin.

---

# Riemann Sum

$f(x) = x^3 - 2x^2 + 2$. Find area under curve from $x=0$ to $x=3$.

--

```{r reimann plot function}
# function to plot Riemann sums

riemann_plot <- function(x_minimum, x_maximum, f, h,
                         rectangle_outline = 'black'){
  
  # create x and y variables
  x <- seq(x_minimum,x_maximum,0.1)
  y <- f(x)
  
  # rectangle dimensions
  x_rectangles <- seq(x_minimum, x_maximum - h, h)
  y_rectangles <- f(seq(x_minimum, x_maximum - h, h))
  
  # approximate area
  approximate_area <- sum(y_rectangles * h)
  
  # rectangle dataframe for plot
  rectangles <- tibble(xmin = x_rectangles,
                       xmax = seq(x_minimum + h, x_maximum, h),
                       ymin = 0,
                       ymax = f(xmin))
  
  
  # return ggplot object
  ggplot() +
    geom_line(data = tibble(x,y),
              aes(x=x,y=y)) +
    geom_rect(data = rectangles,
              aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax),
              fill = 'blue', color = rectangle_outline, alpha = 0.3) +
    labs(title = paste0("h = ", h, 
                        ", Approximate Area = ", round(approximate_area, 3))) +
    theme_bw()
}

```

```{r reimann sum 1}

f <- function(x){
  x^3 - 2*x^2 + 2
}

riemann_plot(x_minimum = 0,
             x_maximum = 3,
             f = f,
             h = 0.5)
```

---

# Riemann Sum

$f(x) = x^3 - 2x^2 + 2$. Find area under curve from $x=0$ to $x=3$.

```{r reimann sum 3}

riemann_plot(x_minimum = 0,
             x_maximum = 3,
             f = f,
             h = 0.25)
```

---

# Riemann Sum

$f(x) = x^3 - 2x^2 + 2$. Find area under curve from $x=0$ to $x=3$.

```{r reimann sum 4}

riemann_plot(x_minimum = 0,
             x_maximum = 3,
             f = f,
             h = 0.1)
```

---

# Riemann Sum

$f(x) = x^3 - 2x^2 + 2$. Find area under curve from $x=0$ to $x=3$.

```{r reimann sum 5}

riemann_plot(x_minimum = 0,
             x_maximum = 3,
             f = f,
             h = 0.01)
```

---

# Riemann Sum

$f(x) = x^3 - 2x^2 + 2$. Find area under curve from $x=0$ to $x=3$.

```{r reimann sum 6}

riemann_plot(x_minimum = 0,
             x_maximum = 3,
             f = f,
             h = 0.001,
             rectangle_outline = NA)
```


---

## Integral Notation

$$\lim_{h \to 0} \sum f(x) \cdot h = \int f(x)dx $$

--

$dx$ is an "infinitesimal" (infinitely small value). 

So $\int f(x)dx$ is the area of an infinite number of infinitely skinny rectangles.

--

If we want the area under a curve between $a$ and $b$, we denote it like so:

$$\int_a^b f(x)dx$$

---

## There has to be an easier way!

What we want is a function $F(x)$; let's call it the **area function**.
  
- $F(a)$ gives the area under $f(x)$ between $-\infty$ and $a$.

- $F(b) - F(a)$ gives the area under $f(x)$ between $a$ and $b$.

--

- As $h$ approaches zero, our skinny rectangles become a better and better approximation of the area function...

$$f(x) \cdot h = \lim_{h \to 0} F(x+h) - F(x)$$

--

Divide by $h$ on both sides:

$$f(x) = \lim_{h \to 0} \frac{F(x+h) - F(x)}{h}$$

--

Hey. We've seen that before!

--

$f(x) = F'(x)$. In other words, $F(x)$ is the *antiderivative* of $f(x)$.


---

# The Fundamental Theorem of Calculus

$$\int_a^b f(x)dx =  F(b) - F(a)$$

--

Computing the area under the curve and taking the antiderivative are equivalent operations!

---

# Try It!

If $f(x) = x$, find the area under the curve between $x=0$ and $x=4$.

```{r plot triangle}

x = -1:5

ggplot(data = tibble(x)) +
  geom_line(aes(x=x,y=x), size = 1) +
  geom_vline(xintercept = 0) +
  geom_hline(yintercept = 0) +
  xlab('x') + ylab('f(x)') +
  geom_vline(xintercept = 4, linetype = 'dashed', color = 'blue')

```

--

**Hint**: It's a triangle, so the answer should be 8.

---

# The Area Under a Curve

If $f(x) = x$, find the area under the curve between $x=0$ and $x=4$.

Use the Fundamental Theorem of Calculus:

$$\int_0^4 f(x)dx = F(4) - F(0)$$
--

$F(x) = \frac{1}{2}x^2 + C$

--

$F(4) - F(0) = \frac{1}{2}\cdot4^2 = 8$

---

## Now a Nonlinear Example...

If $f(x) = x^3 - 2x^2 + 2$, find the area under the curve between $x=0$ and $x=3$.

--

$$\int_0^3 f(x)dx = F(3) - F(0)$$

--

$F(x) = \frac{1}{4}x^4 - \frac{2}{3}x^3 + 2x + C$

--

$F(3) - F(0) = \frac{1}{4}3^4 - \frac{2}{3}3^3 + 2(3) - [\frac{1}{4}0^4 - \frac{2}{3}0^3 + 2(0)]$

--
$= 8.25$

---

## Now a Nonlinear Example...

That's the same answer that we got from the skinny rectangles!

```{r reimann sum 7}

riemann_plot(x_minimum = 0,
             x_maximum = 3,
             f = f,
             h = 0.001,
             rectangle_outline = NA)
```