---
title: "Calculus Essentials"
subtitle: "Integrals and The Fundamental Theorem of Calculus"
format: 
  revealjs:
    incremental: true
editor: visual
echo: false
cache: true
---

## Area Under A Curve

To compute a p-value, we need to know the area under the shaded part of the probability distribution function. We know that the area under the *entire* curve equals 1, but what fraction of that area is shaded below?

```{r}
set.seed(42)
pop <- round(runif(1e5, min = 18, max = 100))

samp <- sample(pop, size = 100)

se <- sqrt(var(samp)/length(samp))

library(tidyverse)

# plot a gaussian and shade a part of it
mu <- 60
sigma <- se

# http://www.sthda.com/english/wiki/ggplot2-area-plot-quick-start-guide-r-software-and-data-visualization
d <- tibble(x = seq(50, 70, 0.01),
            y = dnorm(x, mean = mu, sd = sigma))

p <- ggplot(data = d) +
  geom_line(mapping = aes(x=x,y=y)) +
  geom_area(mapping = aes(x=x, y = if_else(x > mean(samp), y, 0)),
            fill = 'gray', alpha = 0.5) + 
  theme_void() +
  geom_vline(xintercept = mu, linetype = 'dashed')

p
```

## Area Under A Curve

The area of a rectangle is easy. $A = bh$

```{r}
data = tibble(x = -2:2,
              y = 2)

ggplot() +
  geom_rect(aes(xmin = -2, xmax = 2, ymin = 0, ymax = 2),
            alpha = 0.3, fill = 'blue') + 
  geom_line(data = data, aes(x=x,y=y), linewidth = 1) +
  scale_x_continuous(limits = c(-4,4)) +
  scale_y_continuous(limits = c(-4,4)) +
  geom_hline(yintercept = 0) + geom_vline(xintercept = 0) + 
  xlab('x') + ylab('y') + 
  geom_text(aes(x=2.4,y=2, label='f(x)'), size = 5) +
  theme_minimal(base_size = 16)
```

## Area Under A Curve

Curves are harder. We didn't learn this in geometry!

```{r}
f <- function(x){
  x^3 - 2*x^2 + 2
}

# f <- function(x){
#   0.7*x^2
# }

x <- seq(-2,2,0.01)

data <- tibble(x=x,
               y=f(x))

ggplot() +
  geom_line(data = data, aes(x=x,y=y), size = 1) +
  scale_x_continuous(limits = c(-4,4)) +
  scale_y_continuous(limits = c(-4,4)) +
  geom_hline(yintercept = 0) + geom_vline(xintercept = 0) + 
  xlab('x') + ylab('y') + 
  geom_text(aes(x=2.4,y=2, label='f(x)'), size = 5) +
  theme_minimal(base_size = 16)
```

## Area Under A Curve

-   With derivatives, we approximated a hard problem (the slope of a curve) using an easy problem (the slope of a line) by zooming in close enough.
-   With integrals, we'll use a similar trick.
-   We approximate a hard problem (the area under a curve) using an easy problem (the area of a bunch of rectangles) by making the rectangles *really* thin.

## Reimann Sum

$f(x) = x^3 - 2x^2 + 2$. Find area under curve from $x=0$ to $x=3$.

. . .

```{r}
riemann_plot <- function(x_minimum, x_maximum, f, h,
                         rectangle_outline = 'black'){
  
  # create x and y variables
  x <- seq(x_minimum,x_maximum,0.1)
  y <- f(x)
  
  # rectangle dimensions
  x_rectangles <- seq(x_minimum, x_maximum - h, h)
  y_rectangles <- f(seq(x_minimum, x_maximum - h, h))
  
  # approximate area
  approximate_area <- sum(y_rectangles * h)
  
  # rectangle dataframe for plot
  rectangles <- tibble(xmin = x_rectangles,
                       xmax = seq(x_minimum + h, x_maximum, h),
                       ymin = 0,
                       ymax = f(xmin))
  
  
  # return ggplot object
  ggplot() +
    geom_line(data = tibble(x,y),
              aes(x=x,y=y)) +
    geom_rect(data = rectangles,
              aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax),
              fill = 'blue', color = rectangle_outline, alpha = 0.3) +
    labs(title = paste0("h = ", h, 
                        ", Approximate Area = ", round(approximate_area, 3))) +
    theme_bw()
}
```

```{r}
f <- function(x){
  x^3 - 2*x^2 + 2
}

riemann_plot(x_minimum = 0,
             x_maximum = 3,
             f = f,
             h = 0.5)
```

## Reimann Sum

$f(x) = x^3 - 2x^2 + 2$. Find area under curve from $x=0$ to $x=3$.

```{r}
riemann_plot(x_minimum = 0,
             x_maximum = 3,
             f = f,
             h = 0.25)
```

## Reimann Sum

$f(x) = x^3 - 2x^2 + 2$. Find area under curve from $x=0$ to $x=3$.

```{r}
riemann_plot(x_minimum = 0,
             x_maximum = 3,
             f = f,
             h = 0.1)
```

## Reimann Sum

$f(x) = x^3 - 2x^2 + 2$. Find area under curve from $x=0$ to $x=3$.

```{r}
riemann_plot(x_minimum = 0,
             x_maximum = 3,
             f = f,
             h = 0.01)
```

## Reimann Sum

$f(x) = x^3 - 2x^2 + 2$. Find area under curve from $x=0$ to $x=3$.

```{r}
riemann_plot(x_minimum = 0,
             x_maximum = 3,
             f = f,
             h = 0.001,
             rectangle_outline = NA)
```

## Integral Notation

$$\lim_{h \to 0} \sum f(x) \cdot h = \int f(x)dx $$

. . .

$dx$ is an "infinitesimal" (infinitely small value).

So $\int f(x)dx$ is the area of an infinite number of infinitely skinny rectangles.

. . .

If we want the area under a curve between $a$ and $b$, we denote it like so:

$$\int_a^b f(x)dx$$

## The Area Function $F(x)$

. . .

There has to be an easier way to compute the area under a curve than taking the sum of an nearly infinite number of skinny rectangles...

. . .

What we want is a function $F(x)$; let's call it the **area function**.

-   $F(a)$ gives the area under $f(x)$ between $-\infty$ and $a$.

-   $F(b) - F(a)$ gives the area under $f(x)$ between $a$ and $b$.

## The Area Function $F(x)$

As $h$ approaches zero, our skinny rectangles should become a better and better approximation of this area function...

$$f(x) \cdot h = \lim_{h \to 0} F(x+h) - F(x)$$

. . .

Divide on both sides by $h$.

$$f(x) = \lim_{h \to 0} \frac{F(x+h) - F(x)}{h}$$

. . .

Hey doesn't that look familiar?

. . .

$f(x) = F'(x)$. In other words, $F(x)$ is the *antiderivative*.

## The Fundamental Theorem of Calculus

$$\int_a^b f(x)dx =  F(b) - F(a)$$

Finding the area under the curve and taking the antiderivative are equivalent operations!

## Let's Try It!

If $f(x) = x$, find the area under the curve between $x=0$ and $x=4$.

```{r}
x = -1:5

ggplot(data = tibble(x)) +
  geom_line(aes(x=x,y=x), size = 1) +
  geom_vline(xintercept = 0) +
  geom_hline(yintercept = 0) +
  xlab('x') + ylab('f(x)') +
  geom_vline(xintercept = 4, linetype = 'dashed', color = 'blue') +
  theme_minimal(base_size = 16)
```

## Let's Try It!

If $f(x) = x$, find the area under the curve between $x=0$ and $x=4$.

. . .

Use the Fundamental Theorem of Calculus

$$
\int_0^4 f(x)dx = F(4) - F(0)
$$

$F(x) = \frac{1}{2}x^2 + C$

. . .

$F(4) - F(0) = \frac{1}{2}\cdot4^2 = 8$

## Now A Nonlinear Example...

If $f(x) = x^3 - 2x^2 + 2$, find the area under the curve between $x=0$ and $x=3$.

. . .

$$\int_0^3 f(x)dx = F(3) - F(0)$$

. . .

$F(x) = \frac{1}{4}x^4 - \frac{2}{3}x^3 + 2x + C$

. . .

$F(3) - F(0) = \frac{1}{4}3^4 - \frac{2}{3}3^3 + 2(3) - [\frac{1}{4}0^4 - \frac{2}{3}0^3 + 2(0)]$

. . .

$= 8.25$

## Now A Nonlinear Example...

That's the same answer that we got from the skinny rectangles!

```{r}
riemann_plot(x_minimum = 0,
             x_maximum = 3,
             f = f,
             h = 0.001,
             rectangle_outline = NA)
```
