---
title: "Probability Distributions"
author: "Joe Ornstein"
date: "9/15/2021"
output: 
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

# What is a probability distribution?

A **probability distribution** assigns a number from 0 to 1 to every possible value that a variable can take. All of the values in the distribution must sum to 1.

To illustrate, let's look the CES:

```{r load-ces}
library(tidyverse)
library(here)

load(
  here('data/ces-2020/cleaned-CES.RData')
)
```

## Discrete Variables

Discrete variables can take only a given set of values.

Binary variables are a special type of discrete variable that only takes 2 values.

```{r gender}

ces %>% 
  count(gender) %>% 
  # probability is the total number of people in a category divided by the total number of people in all categories
  mutate(probability = n / sum(n)) 

```

Each probability value tells you - if you were to draw a survey respondent at random - what's the chance they would be female or male?

## Categorical Variables

Categorical variables are a discrete variable that could take multiple values.

```{r race}

ces %>% 
  count(race) %>% 
  mutate(probability = n / sum(n))

```

## Ordinal Variables

An ordinal variable is a categorical variable, but there is some ordering. We know the order, but not the relative distances between the values.

```{r ordinal}

ces %>% 
  filter(!is.na(trump_approval)) %>% 
  count(trump_approval) %>% 
  mutate(probability = n / sum(n))

# plot that
ces %>% 
  filter(!is.na(trump_approval)) %>% 
  count(trump_approval) %>% 
  mutate(probability = n / sum(n)) %>% 
  ggplot(mapping = aes(x = trump_approval, y = probability)) + 
  geom_col()
```

## Continuous Variables

A continuous variable could, in principle, take any numeric value.

```{r age}

# *could* do it this way
ces %>% 
  count(age) %>% 
  mutate(probability = n / sum(n))

# instead of computing the probability in each category, we typically summarize by binning values together
ggplot(data = ces,
       mapping = aes(x = age)) +
  geom_histogram(color = 'black',
                 binwidth = 5)

# instead of a histogram, you could use a density plot
ggplot(data = ces,
       mapping = aes(x = age)) +
  geom_density(alpha = 0.5, fill = 'gray') +
  theme_classic()
```

# Joint Distributions

```{r joint-distribution}

joint_distribution <- ces %>% 
  count(gender, educ) %>% 
  mutate(probability = n / sum(n))

sum(joint_distribution$probability)
```

Joint probability: $P(A, B)$

Probability of observing (in the CES) a female with a 2-year degree: $P(female, 2 year) = 0.0635$.

# Conditional Probability

If joint probability told us the probability of observing someone who is female *and* has a two year degree...

Conditional probability tells us the probability that *if* we observe someone who has a 2-year degree, what's the chance they're a female? Or, conversely, *if* we observe a female, what's the chance she has a two-year degree?

The way we denote conditional probability is with a vertical bar:

$P(A|B)$

And it's related to joint probability this way:

$P(A|B) = P(A, B) / P(B)$

```{r conditional-probability}
# P(educ | gender)
conditional_probability1 <- ces %>% 
  group_by(gender, educ) %>% 
  summarize(freq = n()) %>% 
  mutate(conditional_probability = freq / sum(freq))

conditional_probability1

# P(gender | educ)
conditional_probability2 <- ces %>% 
  group_by(educ, gender) %>% 
  summarize(freq = n()) %>% 
  mutate(conditional_probability = freq / sum(freq))

conditional_probability2
```
