---
title: "Week 7: Probability"
author: Joe Ornstein

echo: true
message: false
warning: false
---

This week, we review the basics of **probability theory**, which will provide us with the tools we need to express how certain or uncertain we are about the quantities of interest that we estimate from data. By the end of this week, you will be able to:

-   Describe the properties of probability distribution functions, starting with two that are particularly ubiquitous in social science: the Bernoulli and Normal distributions.

-   Use the Central Limit Theorem to predict whether a sampling distribution will be normally distributed

-   Compute the area under curves and the expected value of a probability distribution using integral calculus

# Reading

-   **DAFSS** Chapter 6

# Problem Set

In this problem set, you'll demonstrate several key concepts of probability theory by randomly sampling from a deck of cards. Submit your responses as a knitted R script or Quarto document.

1.  Load the [cards.csv](data/cards.csv) dataset. How many cards are there? What percent of the cards are red? What is the average number? What is the variance of the cards' numbers? Plot the `number` variable as a histogram.
2.  Sample 50 cards at random. What percent of this sample is red? What's the average number? (Hint: one approach would be to sample 50 index values and subset those rows of the dataset.)
3.  Using a `for()` loop, let's repeat that process 10,000 times. In each iteration of the loop, sample 50 cards at random and record the average number of the sample as an element in a vector. This is the **sampling distribution**.
4.  Let's make sure that the sample mean is an unbiased estimator of the true value. Verify that the mean of the sampling distribution is approximately equal to the population mean.
5.  Now let's investigate how far away from the truth the sample mean can get. What is the variance of the sampling distribution? In what percent of the samples that you drew is the sample mean off by more than 1 from the population mean?
6.  Plot a histogram of the sampling distribution. What probability distribution does the Central Limit Theorem say should approximately describe these sample means? Overlay that histogram with a curve of the appropriate distribution (see "Drawing Curves over Histograms" in the Class Notes below).
7.  According to the probability distribution function you specified in problem 6, what is the probability that the sample mean will be off by more than 1 from the truth?
8.  **Bonus**. Suppose somebody handed you this deck of cards and asked you to estimate the average number in the deck. If you wanted to be 99.9% certain that your estimated value is no more than 1 away from the true value, how many cards should you sample?

# Class Notes

Over the past few weeks, we have been computing a variety of statistics to try and estimate quantities of interest from our data. We've estimated population averages using sample means. We've estimated average treatment effect using differences-in-means and linear models. For reasons we've discussed, these statistics represent our *best guess* at the truth, given the [fundamental limits of our knowledge](01-introduction.html).

But when communicating the findings of our research, we don't just want to report our best guess. We want to honestly state how certain we are about that guess, given the strength of the evidence we've collected. We want to report not only a single value, but a *range* of the most likely values. To do that, we need to think carefully about the random process that generated our dataset.

## Sampling

The book and the problem set have motivated probability theory with things like coins and cards. This is all very nice for building intuition, but in class, I want us to focus on how all this connects to the concept of the **sampling distribution**.

If we were to randomly select a person from the United States population, what is the probability we would select a female? According to the [2020 Census](https://www2.census.gov/library/publications/decennial/2020/census-briefs/c2020br-06.pdf), here are the total number of males and females in the population.

```{r}
females <- 168.8 * 1e6 # 168.8 million
males <- 162.7 * 1e6 # 162.7 million

probability <- females / (males + females)
probability
```

Let's sample one.

```{r}
#| echo: FALSE
set.seed(1)
```

```{r}
sample(c('F', 'M'), 
       size = 1, 
       prob = c(probability, 1-probability))
```

This is a **Bernoulli** random variable. Two possible outcomes—one with probability $p$ and the other with probability $1-p$. What if we randomly sampled 435 people? How many would be female?

```{r}
my_sample <- sample(c('F', 'M'), 
                    size = 435,
                    replace = TRUE,
                    prob = c(probability, 1-probability))

sum(my_sample == 'F')
```

If you take the sum of a bunch of independent Bernoulli random variables, it follows what's called a **binomial distribution**. Let's do that repeatedly, building up a sampling distribution.

```{r}
sampling_distribution <- c()

for(i in 1:10000){
  my_sample <- sample(c('F', 'M'), 
                    size = 435,
                    replace = TRUE,
                    prob = c(probability, 1-probability))
  sampling_distribution[i] <- sum(my_sample == 'F')
}
```

## Drawing Curves over Histograms

I don't think it's explained in the book how the authors plot those nice curves of the distribution functions, so I wanted to include an example here.

```{r}
# generate a bunch of random normal draws
Z <- rnorm(1e4, mean = 2, sd = 2)
hist(Z, freq = FALSE)
# add curve
x <- seq(min(Z), max(Z), 0.1)
lines( x = x,
       y = dnorm(x, mean = 2, sd = 2),
       lty = 'dashed' )
```

Now let's create that same plot for the sampling distribution we just simulated, which should follow the binomial distribution.

```{r}
hist(sampling_distribution, freq = FALSE)
x <- seq(min(sampling_distribution), max(sampling_distribution), by = 1)
lines(x = x,
      y = dbinom(x,
                 size = 435,
                 prob = probability),
      lty = 'dashed')
```

Using this sampling distribution, we can explore a few key ideas.

## Expected Value

On average, how many females should we expect to be in our sample? Well, the expected value of a binomial distribution is $np$, where $n$ is the sample size and $p$ is the probability that each person sampled will be a female.

```{r}
expected_value <- 435 * probability

expected_value
```

We say that our sample statistic is an **unbiased** estimator if its average value is roughly equal to that expected value.

```{r}
mean(sampling_distribution)
```

I want to take a moment to marvel at this fact before we move on. There are over 330 million people in the United States. If I want to know how many of them are females, I could do one of two things. Either I could try to contact every single person in the country—as the Census Bureau does every 10 years—and count up the number of males and females. Or I could take a random sample of a few hundred people, and the percent female in that sample will, on average, give us a good approximation of what the other 330 million look like.

But [averages can be deceiving](https://www.reddit.com/r/Jokes/comments/1fbepm/a_physicist_a_biologist_and_a_statistician_go/)! We don't just want to know whether we'll be right *on average*. We want to know how far from the truth we might end up with our particular sample.

## Standard Error

What is the standard deviation of the sampling distribution?

```{r}
sd(sampling_distribution)
```

The standard deviation of a sampling distribution is is called the **standard error**, and it's a very important number for understanding hypothesis tests. The larger the standard error, the wider the range of sample statistics that could have been computed from our population, and the less certain we should be about our particular value.

## The Central Limit Theorem

Next, notice the *shape* of the sampling distribution in the figure above. It's this nice bell curve that's well-approximated by the "normal" distribution . Why is the sampling distribution normally shaped, even though the thing we're sampling from is not normal distributed? Well that's one of the most fascinating and magical theorems in all of statistics: the **Central Limit Theorem**. The sampling distribution of the mean is approximately normally distributed—as long as you have a sufficiently large sample size.

Intuition: some samples will, by chance, contain an unrepresentatively large number of females. Some will contain an unrepresentatively *small* number females. But most of the time, the females and non-females will cancel each other out, such that the mass of the distribution is centered around the truth.

There are a bunch of nice [simulations](https://seeing-theory.brown.edu/probability-distributions/index.html#section3) that showing the Central Limit Theorem in action. In class, you'll get to play around with one.

Normal distributions are nice. They allow us to say precisely how far a sample estimate is likely to deviate from the truth. For example, we know that about 68% of a normal distribution falls within 1 standard deviation of the mean.

```{r}
expected_value <- mean(sampling_distribution)
standard_error <- sd(sampling_distribution)
num_draws <- length(sampling_distribution)

within_1sd <- sum(sampling_distribution > expected_value - standard_error &
      sampling_distribution < expected_value + standard_error)

within_1sd / num_draws
```

And about 95% of observations will fall within two standard deviations.

```{r}
within_2sd <- sum(sampling_distribution > expected_value - 2*standard_error &
                        sampling_distribution < expected_value + 2*standard_error)

within_2sd / num_draws
```

And about 99.7% will fall within 3 standard deviations. It's super rare to get an observation that far away from the expected value.

```{r}
within_3sd <- sum(sampling_distribution > expected_value - 3*standard_error &
                        sampling_distribution < expected_value + 3*standard_error)

within_3sd / num_draws
```

As of July 2024, there are 125 females in the 435-member US House of Representatives. What's the probability that we'd have that few females in a group of 435 people randomly selected from the US population?

```{r}
sum(sampling_distribution <=125)
```

Approximately zero. Out of 10,000 random samples, we didn't get a single one with that few females. To be precise, the probability of a sample statistic that low is 0.0000000000000000005%.

```{r}
pbinom(125, size = 435, prob = probability)
```

## Margin of Error

So if I draw a sample of 435 individuals, then 95% of the time, my sample statistic will be within...

```{r}
2*sd(sampling_distribution)
```

...of the population average. This is what pollsters call the **margin of error**.

Another way of describing this sampling variability is with a **confidence interval**. You construct a confidence interval by taking your sample estimate plus or minus 2 standard errors. Because the sample mean is within 2 standard deviations of the truth 95% of the time, the 95% confidence interval will contain the true value approximately 95% of the time.

## Sampling Distribution of the Difference-In-Means

In class, we'll explore how the estimated effect of a treatment can vary even if we have data on the entire population. For this exercise, load the [potential-outcomes.csv](data/potential-outcomes.csv) dataset.

```{r}
#| echo: FALSE
# create potential outcomes dataset
set.seed(42)
n <- 1000
y0 <- rnorm(n, 10, 5)
y1 <- y0 + rnorm(n, 0, 1)

d <- data.frame(y0,y1)
write.csv(d, 'data/potential-outcomes.csv')

#treated <- sample(c(0,1), size = length(y0), replace = TRUE)
#y <- ifelse(treated==1,y1,y0)

#lm(y~treated)
```

# Additional Resources

-   [Teacup Giraffes](https://tinystats.github.io/teacups-giraffes-and-statistics/06_standardError.html): Module 6 (Standard Error)
