---
title: "Week 1"
description: |
  Tell stories with data
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

## Readings

- [Teacup Giraffes](https://tinystats.github.io/teacups-giraffes-and-statistics/index.html) (Intro to R Module)
- [Healy Preface](https://socviz.co/index.html#preface) (and complete the package installation instructions at the bottom)
- [Healy Chapter 2](https://socviz.co/gettingstarted.html)
- [R4DS Chapter 1](https://r4ds.had.co.nz/introduction.html)

## Overview

This week, we will familiarize ourselves with a suite of software that allows us to learn lessons from data in a transparent, reproducible fashion. Make sure you've completed everything from the [Getting Started](index.html#getting-started) section on the home page.

```{r slides}
knitr::include_url('slides/01-stories.html')
```

## Introductions

On our first day, we will be collecting and analyzing some data about the class itself. Please complete the following survey.[^survey]

[^survey]: If the embedded survey isn't working for some reason, click [here](https://docs.google.com/forms/d/e/1FAIpQLScvfmmeiPeg4Qudq4E7P-Qiyb7h8mF0CrfusOns4hTGoTdOyQ/viewform?usp=sf_link).

```{r survey}
knitr::include_url('https://docs.google.com/forms/d/e/1FAIpQLScvfmmeiPeg4Qudq4E7P-Qiyb7h8mF0CrfusOns4hTGoTdOyQ/viewform?usp=sf_link')

# The results will be available [here](https://docs.google.com/spreadsheets/d/1BTVFoZQjRNB63Pyg_z-hjGICpP_vlcokRy0-SSGX-10/).
```


## Problem Set

Our first problem set will ensure that all the software is functioning on your machine. If you can complete all of the following steps, then you're in good shape!

1. Download the folder from the class GitHub repository and open `intro-to-political-methodology.Rproj`.

2. Open `problem-sets/problem-set-01.R`.

3. Update the author and date in the metadata at the top. (You're the author. The date is today.)

4. Update the script so that it filters the data for your name and sex (or some other name and sex of your choice) instead of the male "Joseph"s.

5. Make sure you have the `tinytex` package installed. This is the library you need to convert RMarkdown documents into PDFs. To install, enter these two lines into your RStudio console.

```{r install-tinytex, eval = FALSE, echo = TRUE}
install.packages('tinytex')
tinytex::install_tinytex()
```

6. Click 'Compile Report' (Ctrl + Shift + K) and output to PDF.

7. Submit the resulting PDF document to eLC at least 24 hours before our next class.[^1]

[^1]: All problem sets are due by the *following* week's class (so this one is due before our second class). Feel free to work together on problem sets, but I encourage you to type out the solutions yourself instead of copying-pasting from others. You can't learn this stuff without doing it. 