---
title: "Problem Set 1"
format: pdf
---

In this problem set, we look at demographic data from Sweden, Kenya, and the *world*, to compare birth and death rates over a 50 year period.

## Crude Birth Rates

First, let's compute crude death rates for each country and period.

```{r}
# read in the data
kenya <- read.csv('INTRO/Kenya.csv')
sweden <- read.csv('INTRO/Sweden.csv')
world <- read.csv('INTRO/World.csv')

# create a function to compute crude birth rates
calculate_cbr <- function(dataset, period){

  # total person-years column
  dataset$py <- dataset$py.women + dataset$py.men

  # compute total person-years and births in the given period
  py_total_period <- sum(dataset$py[dataset$period == period])
  births_total_period <- sum(dataset$births[dataset$period == period])

  # return crude birth rate
  return(births_total_period / py_total_period)

}

calculate_cbr(kenya, '1950-1955')
calculate_cbr(kenya, '2005-2010')

calculate_cbr(sweden, '1950-1955')
calculate_cbr(sweden, '2005-2010')

calculate_cbr(world, '1950-1955')
calculate_cbr(world, '2005-2010')

```

## Age-Specific Fertility Rates

Next, let's compute age-specific fertility rates (ASFR) for women aged 15-49.

```{r}
asfr <- function(dataset, period){

  # just keep the women 15-50 in the period of interest
  dataset <- dataset[dataset$period == period & dataset$age %in% c('15-19','20-24', '25-29', '30-34', '35-39', '40-44', '45-49') ,]

  asfr <- dataset$births / dataset$py.women
  names(asfr) <- dataset$age

  return(asfr)

}

asfr(dataset = kenya, period = '1950-1955')
asfr(dataset = kenya, period = '2005-2010')

asfr(dataset = sweden, period = '1950-1955')
asfr(dataset = sweden, period = '2005-2010')

# that's interesting. Sweden birth rates declined in every age group *except* women in their 30s. That makes sense!

asfr(dataset = world, period = '1950-1955')
asfr(dataset = world, period = '2005-2010')

```

## Total Fertility Rate

Now let's use the ASFRs to compute the total fertility rate (TFR) for each country and period.

```{r}
tfr <- function(dataset, period){

  asfr <- asfr(dataset, period)

  # total fertility rate is each asfr times 5, added together
  sum(asfr * 5)

}

tfr(kenya, '1950-1955')
tfr(kenya, '2005-2010')

tfr(sweden, '1950-1955')
tfr(sweden, '2005-2010')

tfr(world, '1950-1955')
tfr(world, '2005-2010')

```

## Crude Death Rate

Next, we'll compute crude death rates (CDR) for each country and period.

```{r}
calculate_cdr <- function(dataset, period){

  # total person-years column
  dataset$py <- dataset$py.women + dataset$py.men

  # compute total person-years and deaths in the given period
  py_total_period <- sum(dataset$py[dataset$period == period])
  deaths_total_period <- sum(dataset$deaths[dataset$period == period])

  # return crude death rate
  return(deaths_total_period / py_total_period)

}

calculate_cdr(kenya, '1950-1955')
calculate_cdr(kenya, '2005-2010')

calculate_cdr(sweden, '1950-1955')
calculate_cdr(sweden, '2005-2010')

# interesting note here. crude death rates appear to be equal in Kenya and Sweden in the 2005-2010 period!

calculate_cdr(world, '1950-1955')
calculate_cdr(world, '2005-2010')

```

## Age-Specific Death Rate

Now let's compute age-specific death rates (ASDR) for each country and period.

```{r}
asdr <- function(dataset, period){

  # just keep the period of interest
  dataset <- dataset[dataset$period == period,]

  asdr <- dataset$deaths / (dataset$py.women + dataset$py.men)
  names(asdr) <- dataset$age

  return(asdr)

}

asdr(kenya, '1950-1955')
asdr(kenya, '2005-2010')

asdr(sweden, '1950-1955')
asdr(sweden, '2005-2010')

asdr(world, '1950-1955')
asdr(world, '2005-2010')

```

## Counterfactual CDR

Finally, what would Kenya's Crude Death Rate be if it had the age-distribution of Sweden?

```{r}
asdr(kenya, '2005-2010')

sweden$py <- sweden$py.men + sweden$py.women
sweden_population <- sum(sweden$py[sweden$period=='2005-2010'])
sweden_period <- sweden[sweden$period == '2005-2010',]
# the proportion column will now tell me the percent of Swedes in each age bucket during the 2005-2010 period
sweden_period$proportion <- sweden_period$py / sweden_population

# counterfactual CDR is sum(asdr x sweden's population proportions)
sum(asdr(kenya, '2005-2010') * sweden_period$proportion)

```