---
title: "Probability Theory"
# author: "Joseph T. Ornstein"
# date: "June 12, 2020"
output:
  xaringan::moon_reader:
    nature:
      highlightStyle: github
      countIncrementalSlides: false
subtitle: 'POLS 7012: Introduction to Political Methodology'
institute: University of Georgia
---

<style>

.center-middle {
  margin: 0;
  position: absolute;
  top: 50%;
  left: 50%;
  -ms-transform: translate(-50%, -50%);
  transform: translate(-50%, -50%);
}

</style>

```{r Setup, include=FALSE}
library(tidyverse)
theme_set(theme_minimal(base_size = 16))
knitr::opts_chunk$set(comment=NA, fig.width=7, fig.height=5, 
                      fig.align = 'center', out.width = 600,
                      message=FALSE, warning=FALSE, echo=TRUE)
set.seed(42)
```


A lot of probability/statistics courses are pretty abstract; the great thing about learning `R` first is that we can introduce this in a really hands-on way! 

We'll start by learning some functions to draw random values in `R`.

Then Law of Large Numbers, Bayes Rule, Central Limit Theorem

Random Variables and their properties: Expectation, Variance, Covariance


---


## Generating Random Numbers in `R`

The `sample` function draws values at random from a vector:

```{r sample}
x <- 1:20

sample(x, size = 1)

sample(x, size = 1)

sample(x, size = 100, replace = TRUE)
```

---

## Generating Random Numbers in `R`

**Exercise**: Draw 5 random letters. (Note: the vector `LETTERS` is built-in to R.)

--

```{r sample exercise}
sample(LETTERS, size = 5, replace = TRUE)
```

---

## Generating Random Numbers


---

## Creating Your Own Functions

```{r create function}
draw_n_letters <- function(num_letters = 5){
  sample(LETTERS, size = num_letters, replace = TRUE)
}

draw_n_letters(num_letters = 10)

count_Q <- function(num_letters = 5){
  sample_letters <- LETTERS %>% 
    sample(size = num_letters, replace = TRUE)
  
  length(sample_letters[sample_letters == 'Q'])
}

count_Q(100)
```

---

## Repeating Commands with `replicate`

TODO: Questioning lifecycle 

```{r replicate}
replicate(5, draw_n_letters(10))
```

---

## Monte Carlo

Draw a bunch of samples and compute statistic:

```{r monte carlo, out.wdith = '50%'}
Q <- replicate(1000, count_Q(1000))

hist(Q)
```