---
title: 'Problem Set 7: Linear Regression'
author: "Joe Ornstein"
date: "Due November 11, 2020"
output:
  pdf_document: default
  html_document: default
urlcolor: blue
---

```{r setup, include=FALSE}
# For problem set, echo and eval = FALSE
# For answer key, echo and eval = TRUE
knitr::opts_chunk$set(echo = TRUE, eval = TRUE, message = FALSE, warning = FALSE)
```

Create an R project, an R script, load tidyverse, and import the modified Freedom In The World dataset you created for the midterm.

```{r load FIW}
library(tidyverse)
library(readxl)

# load Freedom in the World datset
FIW <- read_xlsx(path = 'data/2020_All_Data_FIW_2013-2020.xlsx', skip = 1, sheet = 'FIW13-20') %>% 
  filter(`C/T` == 'c',
         Edition == 2020) %>% 
  mutate(democracy_score = A1 + A2 + A3 + B1 + B2 + B3 + B4 + C1 + C2 + C3,
         liberty_score = D1 + D2 + D3 + D4 + E1 + E2 + E3 + F1 + F2 + F3 + F4 + G1 + G2 + G3 + G4)
```

## Line of Best Fit

1. Estimate a line of best fit with `liberty_score` as the outcome and `democracy_score` as the explanatory variable.

```{r line of best fit}
lm1 <- lm(liberty_score ~ democracy_score, data = FIW)
```

2. Visualize the line of best fit with `geom_smooth(method = 'lm')`.
  
```{r scatter plot}
ggplot(data = FIW) +
  geom_point(mapping = aes(x=democracy_score, y=liberty_score)) +
  geom_smooth(mapping = aes(x=democracy_score, y=liberty_score),
              method = 'lm') +
  labs(x='Democracy Score', y='Liberty Score') +
  theme_bw()
```
  
3. What is the slope of the relationship? What is the 95% confidence interval around that slope?

```{r slope and confidence intervals}
coef(lm1)['democracy_score']
confint(lm1)['democracy_score',]
```
  
4. Are there any outliers with higher or lower liberty scores than you would expect given their democracy score? Which countries are they? (Hint: there is a vector called `residuals` in the `lm` object you just created. Add it to your FIW dataframe).

```{r residuals}
FIW <- FIW %>% 
  mutate(lm1_residual = lm1$residuals)

# sort by residuals and display the top 5
FIW %>% 
  arrange(-lm1_residual) %>% 
  select(`Country/Territory`, democracy_score, liberty_score, lm1_residual) %>% 
  head(5)
```
Monaco, Benin, Thailand, Eswatini, and Liechtenstein are all more free than we would expect given their level of democracy. (BTW: I had no idea Swaziland was officially renamed in 2018.)

```{r residuals 2}
FIW %>% 
  arrange(lm1_residual) %>% 
  select(`Country/Territory`, democracy_score, liberty_score, lm1_residual) %>% 
  head(5)
```
Iraq, Bhutan, Myanmar, Indonesia, and El Salvador are all *less* free than we would expect given their level of democracy.

## Data Wrangling and OLS
  
I'm including another dataset that I pulled from the World Bank with GDP per capita figures for each country since 1960. Man, it's a mess. We're going to have to tidy it up.

5. Read the dataset into `R` and pivot the data so each row represents a country-year.

6. Keep the most recent year of gdp per capita

7. Create a new variable, `log_gdp_per_capita`, equal to the logarithm of GDP per capita.

```{r read-pivot-filter-and-mutate}
WB <- read_csv('data/API_NY.GDP.PCAP.PP.CD_DS2_en_csv_v2_1495136.csv',
               skip = 4) %>% 
  select(-c(`Indicator Name`, `Indicator Code`, X66)) %>% 
  pivot_longer(cols = `1960`:`2020`,
               names_to = 'year', 
               values_to = 'gdp_per_capita') %>% 
  filter(year == 2019) %>% 
  mutate(log_gdp_per_capita = log(gdp_per_capita))
```

8. Save your cleaned up World Bank dataset to the `data/` folder.

```{r save}
write_csv(WB, 'data/cleaned_WB.csv')
```

9. Merge your cleaned up World Bank dataset with the FIW dataset.

```{r merge with FIW}
# make sure the key variables have the same name
WB <- WB %>% 
  rename(country_name = `Country Name`)

# merge with left_join
data <- FIW %>% 
  rename(country_name = `Country/Territory`) %>% 
  left_join(WB, by = 'country_name')

# Which countries failed to merge?
data %>% 
  filter(is.na(`Country Code`)) %>% 
  pull(country_name)

# Rename those countries in the WB dataset
WB <- WB %>% 
  mutate(country_name = case_when(country_name == 'Bahamas, The' ~ 'Bahamas',
                                  country_name == 'Brunei Darussalam' ~ 'Brunei',
                                  country_name == 'Congo, Rep.' ~ 'Congo (Brazzaville)',
                                  country_name == 'Congo, Dem. Rep.' ~ 'Congo (Kinshasa)',
                                  country_name == 'Egypt, Arab Rep.' ~ 'Egypt',
                                  country_name == 'Iran, Islamic Rep.' ~ 'Iran',
                                  country_name == 'Kyrgyz Republic' ~ 'Kyrgyzstan',
                                  country_name == 'Lao PDR' ~ 'Laos',
                                  country_name == 'Micronesia, Fed. Sts.' ~ 'Micronesia',
                                  country_name == 'Korea, Dem. People’s Rep.' ~ 'North Korea',
                                  country_name == 'Russian Federation' ~ 'Russia',
                                  country_name == 'Slovak Republic' ~ 'Slovakia',
                                  country_name == 'Korea, Rep.' ~ 'South Korea',
                                  country_name == 'Syrian Arab Republic' ~ 'Syria',
                                  country_name == 'Gambia, The' ~ 'The Gambia',
                                  country_name == 'Venezuela, RB' ~ 'Venezuela',
                                  country_name == 'Yemen, Rep.' ~ 'Yemen',
                                  TRUE ~ country_name))

# Try the merge again
data <- FIW %>% 
  rename(country_name = `Country/Territory`) %>% 
  left_join(WB, by = 'country_name')

# Now only Taiwan fails to merge 
# (there's no separate entry in the World Bank Data)
data %>% 
  filter(is.na(`Country Code`)) %>% 
  pull(country_name)
```

10. Estimate a multivariable linear model with `liberty_score` as the outcome variable and `democracy_score` and `log_gdp_per_capita` as the explanatory variables. What are the slope coefficients and 95% confidence intervals?

```{r lm2}
lm2 <- lm(liberty_score ~ democracy_score + log_gdp_per_capita, 
          data = data)

summary(lm2)
```

The estimated slope coefficient for `democracy_score` is a bit smaller than it was before, but not by much. Democracies tend to be wealthier on average, and wealthier places tend to have higher Freedom House scores, but it is not enough to fully explain the relationship we found between democracy and liberty in the midterm.

\newpage

## Why Log GDP Per Capita?

```{r gdp and liberty}
# GDP per capita is right-skewed
ggplot(data) +
  geom_histogram(aes(x=gdp_per_capita), color = 'black')

# On a log scale it's a bit more normal
ggplot(data) +
  geom_histogram(aes(x=gdp_per_capita), color = 'black') +
  scale_x_log10()

# Not a linear relationship!
ggplot(data) + 
  geom_text(aes(x=gdp_per_capita, y=liberty_score, 
                label=`Country Code`), size = 2) +
  labs(x='Log GDP Per Capita',y='Liberty Score') +
  theme_bw()

# This is more plausibly linear
ggplot(data) + 
  geom_text(aes(x=gdp_per_capita, y=liberty_score, 
                label=`Country Code`), size = 2) +
  scale_x_log10() +
  geom_smooth(aes(x=gdp_per_capita, y = liberty_score),
              method = 'lm') +
  labs(x='Log GDP Per Capita',y='Liberty Score') +
  theme_bw()
```