---
title: "Perception of Economy Strength by Education Level"
author: "Sidney Shank"
date: "9/1/2021"
output: pdf_document
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(stringr)
load("data/ces-2020/cleaned-CES.RData")
ces <- ces %>%
  filter(national_economy != "NA") %>%
  mutate(educ = factor(educ, levels = c("No HS", "High school graduate", "Some college",
                                        "2-year", "4-year", "Post-grad")))
```

## Introduction

For this week's problem set, I built off of my graph from last week exploring the relationship between the highest completed education levels of U.S. citizens and their perceptions of the U.S. economy. The data for this analysis was generated by the Cooperative Election Study (CES) across two waves of surveys in late 2020: from September 29 to November 2 and from November 8 to December 14. I focused on responses to the question, "Would you say that OVER THE PAST YEAR the nation's economy has (1) gotten much better; (2) Gotten somewhat better; (3) Stayed about the same; (4) Gotten somewhat worse; (5) Gotten much worse; (6) Not sure" [@schaffner2021]. I compared the responses to this question to those same respondents' self-reported answers to their highest completed level of education, ranging from those who never finished high school to those who hold at least one graduate degree, to produce the graph below.

```{r graph, echo=FALSE}
p <- ggplot(data = ces, mapping = aes(x = national_economy, fill = educ))
p + geom_bar(position = "dodge", mapping = aes(y = ..prop.., group = educ)) +
  labs(y = "Proportion of Responses", x = "Strength of Economy",
       title = "Perception of Economy Strength by Education Level",
       subtitle = "U.S. Perception of Economic Change from Fall/Winter 2019
       to Fall/Winter 2020",
       caption = "Data source: 2020 Cooperative Election Study.") +
  scale_x_discrete(labels = function(x) str_wrap(x, width = 10),
                   limits = c("Gotten much better", "Gotten somewhat better",
                              "Stayed about the same", "Gotten somewhat worse",
                              "Gotten much worse", "Not sure")) +
  scale_fill_discrete(name = "Highest Education Level",
                      labels = c("No high school", "High school diploma", "Some college",
                                 "Two-year degree", "Four-year degree", "Graduate degree"))
```

## Analysis

In general, respondents felt that the economy had "gotten somewhat worse" or "gotten much worse" from the fall of 2019 to the fall of 2020. The most obvious explanation for this shift is the COVID-19 pandemic, which increased unemployment, poverty, and the threat of eviction or homelessness. Though the pandemic likely accounts for most of the uniformity in responses, a few interesting education-level trends do appear.

One clear trend from the graph is that respondents with a graduate degree felt most strongly that the economy had "gotten much worse." In addition to having the highest proportion of respondents in the "much worse" category, post-graduates also had the lowest proportion of responses across every other category compared to all other education levels. Despite these responses, post-graduates were likely the least affected of the surveyed respondents economic downturn during the pandemic. Post-graduate jobs are typically higher-salary and increasingly technological, so they are likely the people best able to weather the pandemic recession, either through continuing their regular job remotely or relying on their proportionally-larger savings. That post-graduates responded to the question with such a strong negative emphasis implies that they likely evaluated this question in terms of the overall health of the economy regardless of their particular economic circumstances. Perhaps these responses make more sense when considering that liberal social and economic views are to some extent correlated with education levels, so the general liberal disagreement with Trump's response to the pandemic may pull those post-graduate responses further negative.

A second observation, one that is less relevant but equally interesting, is the relative proportions of "Not sure" respondents. Respondents who did not graduate high school were proportionally twice as likely or more to choose this response than respondents in any other education level. Additionally, as education levels increase, responses in this category decrease across the board. One theory for this shift across education levels is that higher education levels increases response confidence by giving citizens a greater understanding of the interrelationship between politics and economics, allowing them to develop more thorough arguments to back their position. Another possibility is that as education level rises, the (real or perceived) education gap between survey participants and survey researchers decreases, influencing participants to respond to the survey as though they are answering to their intellectual peers rather than to authority figures. Though outside the scope of this problem set, the larger implications of this second theory on the validity of public opinion surveys and the ways in which education levels between participants and researchers could skew responses are serious and deserve future study (though given that public polling is not an area of interest for me, it's possible that these questions have already been given the attention they deserve).

## References
