---
title: '''Country preferences: Freedom vs. Security'
author: "Ashley Pieper"
date: "9/7/2021"
output: pdf_document
urlcolor: blue
bibliography: references.bib
---

```{r setup, include=FALSE}
# this line tells RMarkdown not to include code or warning messages in the final PDF
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

#load libraries here
library(tidyverse)
library(here)
library(knitr)

```

# Introduction

Question Q150 on the **World Values Survey Wave 7 (2017-2020)** asks respondents, "Most people consider both freedom and security to be important, but if you had to choose between them, which one would you consider more important?"[^1] People who favor freedom over security likely enjoy a relatively high degree of political and social privilege, as it implies that they experience fewer threats to their immediate security. Alternatively, people whose physical safety is constantly threatened due to violence, resource scarcity, natural disasters, etc. are more likely to perceive security as a more immediate need and personal freedom as a luxury. For example, people who have the luxury of choosing to forgo vaccines, mask wearing, and/or other preventative measures during the COVID-19 pandemic likely enjoy greater levels of privilege, and therefore they perceive threats to their personal freedom as more severe than threats to their physical health. For this reason, they may be less likely to support legislation mandating vaccines, masks, and/or other preventative measures. Alternatively, people in less-privileged positions may be more likely to support such legislation as they would be more likely to fear the threat of COVID-19 to their physical, financial, and or job-related security. Therefore, I hypothesize that Western, advanced democracies are more likely to favor freedom over security.

[^1]: @haerpfer2021g

```{r load-WVS}
load(
  here('data', 'raw data', 'WVS_Cross-National_Wave_7_R_v2_0.rdata')
)
```

\newpage

## Country Preferences (Freedom or security - Which is more important?)

```{r figure1, fig.width=12, fig.height=8, fig.cap='Preferences by Country, World Values Survey (2021)'}
`WVS_Cross-National_Wave_7_v2_0` %>% 
  select(country = C_COW_ALPHA,
         freedom_and_security = Q150,
         birthyear = Q261) %>% 
  filter(birthyear >= 1945) %>% 
  mutate(freedom_and_security = case_when(freedom_and_security == 1 ~ 'Freedom',
                                          freedom_and_security == 2 ~ 'Security')) %>% 
  na.omit %>% 
  ggplot(mapping = aes(x = freedom_and_security, 
                       fill = freedom_and_security)) +
  geom_bar() +
  facet_wrap(~country) +
  labs(x = 'Freedom vs. Security', y= 'Value')

```

## Results

The results in **figure 1** show that the overwhelming majority of countries value security over freedom. The one outlier is the United States. This may be in part due to the great degree of privilege Americans experience (all else equal) when compared to other countries. The US Constitution which largely emphasizes individual freedoms over security may also be a contributing factor. This suggests that on average, compared to other countries, Americans may be more likely to oppose legislation related to cyber-security/surveillance than people in other countries. Similarly, this may explain why policies related to firearm restrictions are less popular in the US than in other countries.

# References
