---
title: Urban v. Rural Attitudes Towards Politics in Latin America
author: 
  - Aidan N. González^[University of Georgia]
date: September 7, 2021
output:
  pdf_document:
    number_sections: yes
urlcolor: blue
abstract: Within Latin America, like much of the global south, there is a significant societal gap between urban and rural populations. The common result is a disparity in political engagement and political interest. However, due to the populist movements of the 20th century, Latin America's rural population is far from apolitical. I find the results to be slightly signifcant in that, as a result of historical context, a Latin American individual's location has no effect on the value they give to politics.
bibliography: references.bib
---

```{r setup, include=FALSE}
# this line tells RMarkdown not to include code or warning messages in the final PDF.
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

# load libraries here
library(tidyverse)
library(here)
library(knitr)
```

\newpage

# Introduction

Political sophistication and engagement has long been a topic of discussion, with the consequences of this discussion having serious effects on the function of democracy. The legitimacy of a democracy is dependent on a number of factors with political engagement of the public and faith in government being two major factors. However, within a democracy, faith in government/government legitimacy is closely related to the engagement of the public.

Political engagement, at the risk of oversimplifying an incredibly complex topic, is not homogeneous among all parts of the public. The widely accepted reasoning for disparities in engagement, within North America, is the disparity in political sophistication or ability to understand the nuances of politics. Variance in political sophistication, within the United States, can largely be attributed to a number of socioeconomic factors such as education level, income, and access to high-quality information. While the nature of the factors has changed in the 21st century, with the proliferation of the internet allowing for mass consumption of information of varying quality, income remains a barrier towards political engagement [@zaller1992].

Put bluntly, individuals who are focused on subsistence likely do not have the time to participate in politics. However, the focus of this analysis is not to measure the level of political participation in Latin America, but to measure the value individuals place on politics. Latin America is a unique case in the study of value placed on politics, given their history of populist leaders and the persisting high rate of poverty @piñeiro2016a Elite discourse, many would argue, shapes the public perception of issues and the public's perception of whether an issue is important [@jacoby1995]. Latin America's string of populist leader, who rose to prominence as a result of the impoverished mass's disatisfaction with their conditions, resulted in this political interest disparity, seen in parts of North America, to be diminished.

The masses of Latin America, namely those living in rural areas far removed from the political elite and governmental bases of operation, i.e. capital cities, likely still see the value of politics and the importance of government policies. Latin America's history of populist leaders who motivated the poor/rural masses and given the fact governmental policies have authority over agriculture, the main industry rural individuals depend on, I hypothesize geographic location will have no effect on the importance one gives to politics.

```{r load-wvs}
load(here("~/IntroPolMethods/intro-political-methodology/WVS_Cross-National_Wave_7_R_v2_0 2.rdata"))
```

# Results

```{r figure1, fig.cap='Differences between Rural and Urban Latin America, World Value Survey (2020)'}
library(tidyverse)
`WVS_Cross-National_Wave_7_v2_0` %>%
  select(country=C_COW_ALPHA,
         rural_urban=H_URBRURAL,
         importance_of_politics=Q4)%>%
  filter(rural_urban>=1)%>%
  #Sort urban and rural into the two categories
  mutate(rural_or_urban=case_when(rural_urban<2~'Urban',
                                  rural_urban<3~'Rural'))%>%
  #Combine to create a Latin America column. Then pipe into ggplot
  mutate(latin_america='MEX', 'BOL', 'COL', 'CHL','BRA',
         'GUA', 'ECU', 'NIC')%>%
  ggplot(mapping=aes(x=importance_of_politics))+
  geom_density()+
  facet_wrap(~rural_or_urban)+
  theme_minimal()+
  theme(panel.grid.minor=element_blank())+
  labs(title="Importance of Politics to Rural and Urban Latin America",x="Importance of Politics",
       y='Density of Responses')

```

\`\`\`

# Discussion

While the figure above only allows for an estimation of difference, a visual estimate in particular, there is a clear lack of significant difference. While those living in Urban areas of Latin America did, on average, give democracy a slightly rating of importance, those living in Rural areas followed a similar trend. This figure tentatively confirms my hypothesis as there was no significant difference based on geography. Future studies should look to see what other differences in Latin American culture contribute to this consistent value given to politics. Might nations who have not, in the past, gone through a populist movement, have a Rural population who share a similar sentiment. \$\$

# References
