---
title: "Problem Set 1"
author: "Stewart Ulrich"
date: "9/2/2020"
output: html_document
---
```{r}
# Load Data
load('../data/anes_pilot_2019.RData')

#Find number of rows
nrow(data)
```
```{r}
#Find number of columns
ncol(data)
```
```{r}
#Create age variable 
data$age <- 2020-data$birthyr
```

```{r}
#Find median age
median(data$age)
```

```{r}
#Histogram of age variable 
hist(data$age)
```

```{r}
#Table of vote16 variable
table(data$vote16)
```
#603 respondents skipped this question 

```{r}
#Remove NA's
data$vote16[data$vote16 == -1] <- NA
```

```{r}
#Create table without NA's
table(data$vote16)
```
#Removed the -1 respondents as they are now coded as NA

```{r}
#Table of liveurban variable
table(data$liveurban)
```
#Most respondents are in option 3, which is a suburb

```{r}
#2 way table of liveurban and vote20jb
table(data$liveurban, data$vote20jb)
```
#Rural respondents are most likely to vote for Donald Trump 
#Urban respondents are most likely to vote for Joe Biden 

```{r}
#Table of reg1a variable, whether people are registred to vote
table(data$reg1a)
```
#Many skipped this question, but most who did are registered to vote 

```{r}
#Table of reg1b, another question of whether people are registered to vote 
table(data$reg1b)
```
#Same with the other registered to vote question 

```{r}
#Find mean of feeling thermometer for Trump
mean(data$fttrump)
```
#The average feeling thermometer score for Trump 

```{r}
#Find mean of feeling thermometer for Obama
mean(data$ftobama)
```
#The average feeling thermometer score for Obama. Higher than Trump's 






