---
title: Problem Set 3 Answer Key
#date: "`r Sys.Date()`"
authors:
  - name: Joseph T. Ornstein
    email: jornstein@uga.edu
    address: University of Georgia
abstract: |
  In this short paper, I demonstrate how to embed basic data analysis and visualization into an `RMarkdown` document, creating a reproducible workflow.
keywords:
  - RMarkdown
  - reproducible research
  - World Values Survey
#fontsize: 12pt
#spacing: halfline # could also be oneline
#classoptions:
#  - endnotes
bibliography: refs.bib
biblio-style: apsr
output: rticles::oup_article
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE) # By default, hide code; set to TRUE to see code
knitr::opts_chunk$set(fig.pos = 'p') # Places figures on their own pages
knitr::opts_chunk$set(out.width = '100%', dpi=300) # Figure resolution and size
knitr::opts_chunk$set(fig.env="figure") # Latex figure environment

library(xtable) # Creates tables that follow OUP guidelines; other options, such as kable, may also be workable
library(tidyverse)
```


```{r load data}
load('data/World_Values_Survey_Wave_7_Sample.RData')
```

# Introduction

The World Values Survey (WVS) is an extensive survey of cultural values; Wave 7 of the survey interviewed `r nrow(wvs)` people in `r length(unique(wvs$B_COUNTRY))` countries.  It has been cited in papers like @inglehartModernizationCulturalChange2000 and @minkovHofstedeFifthDimension2012.

# Results

Fig. \ref{fig:fig1} plots the number of respondents in each country who say they are willing to fight for their country in the event of another war. Note that the largest outlier is Japan, where a substantial majority say no. 

```{r fig1, fig.cap="Willingness to fight for one's country, by country", fig.width=5, fig.height=7}

# Clean up Q150
wvs <- wvs %>% 
  mutate(fight_for_country = if_else(Q151 == 1, 'Yes', 'No'))

fig1 <- ggplot(data = wvs %>% 
                 filter(!is.na(fight_for_country))) +
  geom_bar(aes(x = fight_for_country)) +
  facet_wrap(~B_COUNTRY_ALPHA) +
  theme_classic() +
  labs(x = 'Willing to Fight for Your Country?',
       y = 'Count')

fig1
```




# References
